---
title: "Lab 11 - Clustering"
subtitle: "ANTH 3900 / 5203"
theme: journal
embed-resources: true
echo: true
format: html
editor_options: 
  chunk_output_type: console
---

***Due: Wednesday 4/22 by 5PM!***

Please complete all components described below.

**To submit:**

1. Clone your Week 14 Repo if you have not yet done so.

2. Create a new Quarto document. Title it "NAME_Lab11_Spring2026". 

3. At the very top of the document under title and author please be sure to put `embed-resources: true` AND `echo: true`. **This is a requirement each week!**

4. Complete all steps below.

5. Render and save. Make sure it is saved in the same folder as the RProject file for Week 14. 

6. Commit and Push.

**Remember:**

1.  Whenever you are using a package. You must open it. Remember `library()`. This must be in a code chunk in your document. 

2.  If you are using data, you must import it inside a code chunk. You cannot just use the import data button in the `Environment` pane. 

3.  Pay attention to your warnings and errors. They mean something.
  - Remember, ggplot requires a `+` sign, not the pipe.

4. Remember to `set.seed()` if necessary.

## Part 1 - K-Means

Please use the `howells` dataset found in the data folder under the assignment. 

1. Import the data

2.  Filter out ***5*** populations. AKA - Select only 5 groups from the `Population` columns. 

3.  Complete a K-Means clustering. 
    * Remember to scale and remove NAs.
    * Use **at least 20** features (i.e., 20 numeric columns).
    * Determine best number of clusters. Justify this determination. 
    * Run the clustering algorithm. 
    * Visualize the results. 
    * Interpret.
  
## Part 2 - K-Mediods

Please use the `BreastCancer` dataset from the mlbench package.

```{r}
#| echo: true
#| eval: false

install.packages("mlbench") # only run once
library(mlbench)
data("BreastCancer")
```

I am interested in whether tumor traits cluster by benign or malignant type. 

1. Calculate the dissimilarity matrix of all variables EXCEPT the `ID` variable and the variable of interest `Class`.
2.  Complete K-Mediods Clustering. 
3. Visualize with `fviz_clust`.
4. Interpret. 

## Part 3 - Hierarchical Clusters

Please use the `carnivora` dataset from the ape package.

```{r}
#| echo: true
#| eval: false

install.packages("ape") # only run once
library(ape)
data("carnivora")
```

1. Select only Column 6 to Column 12 for the analysis. Note, you may use other columns later for interpretation.

2. Remove NAs and Scale Data. 

3. Cluster the data using hierarchical clustering. 
    * Decide on agglomerative vs. divisive. Defend you choice. 
    * Choice a comparison method. 
    * Visualize your results. 
  
4. Can you determine if any clusters correspond to any of the factor classes (i.e., Family or Superfamily)? Note, each leaf will be represented by a number - this number is the row index number.

#

***Don't forget to save, render, commit, and push!***