---
title: "Lab 9 - Generalized Linear Models - ANSWER KEY"
embed-resources: true
theme: journal
toc: true
format: html
editor_options: 
  chunk_output_type: console
---

**Please see below for answers to all parts of Lab 8. Note, I provide you one way to do the code - there are others. That said, all code relates to the topic of the week - linear regression.**

[**Scoring Summary**]{.underline}

* 10 points = submission to Github
  - **You must upload the QMD and HTML for full credit**
* 5 points = Code that runs (no errors, all chunks visible)
* 5 points = Correct answers. (Regardless of code)

# Answers

## Part 1

## Import Data

```{r}
#| echo: true

chimp <- read.csv("~/Documents/assignments_SP2026/Lab9_GLM/data/chimpanzees.csv")
```

### Fit Logistic Regression Models

```{r}
#| echo: true

fit_chimp1 <- glm(pulled_left ~ prosoc_left, data = chimp)
fit_chimp2 <- glm(pulled_left ~ prosoc_left + chose_prosoc + condition, data = chimp)
```

### Compare Models

Students can add any number of predictors to the second model. Interpretation should relate to the model with the **lowest** aic value. In this case, ***Model 1*** is best. All interpretation will be made from model 1. 

```{r}
#| echo: true

fit_chimp1$aic
fit_chimp2$aic
```

### Interpret Best model

The odds of a chimpanzee pulling the lever given their friend was on the other side increases 1.14x. This indicates chimpanzees are likely prosocial.

```{r}
#| echo: true

summary(fit_chimp1)
exp(fit_chimp1$coefficients)
```

## Part 2

### Import Data

```{r}
#| echo: true

pop <- read.csv("~/Documents/assignments_SP2026/Lab9_GLM/data/Kline.csv")
```

### Modify Data

```{r}
#| echo: true
#| message: false

library(tidyverse)
library(magrittr)

pop %<>% mutate(log_pop = log(population), contact_high = ifelse(contact == "high", 1, 0))

```

### Fit Poisson Model

```{r}
#| echo: true

fit_pop1 <- glm(total_tools ~ log_pop + contact_high, data = pop, family = poisson)
```

### Interpretation

In general, total tools increase with an increase in population size BUT there is a decrease in tool count if there is low contact between populations. Said another way, tool number increases if contact increases. 

```{r}
#| echo: true

summary(fit_pop1)
exp(fit_pop1$coefficients)
```

### Compare to Additional Models

Both models are relativley similar. Therefore, I would select the model with **less** predictors (because it is simpler). The interpretation does not change from above. 

```{r}
#| echo: true

fit_pop2 <- glm(total_tools ~ log_pop + contact_high + mean_TU, data = pop, family = poisson)

fit_pop1$aic
fit_pop2$aic
```

## Part 3

### Prepare Data

```{r}
#| echo: true

library(MASS)
data("snails")
```

### Fit a Binomial Model

```{r}
#| echo: true

fit_snail1 <- glm(cbind(Deaths, N - Deaths) ~ Species + Exposure + Rel.Hum + Temp, data = snails, family = binomial)
```

### Interpretation

Species B is more likely to die in comparison to species A. Exposure increases risk of death 4-fold. A decrease in relative humidity decreases risk of death. Temperature has a minimal if not slightly positive effect.

```{r}
#| echo: true

summary(fit_snail1)
exp(fit_snail1$coefficients)
```

#