---
title: "Lab 2 - Spring 2026"
subtitle: "ANTH 3900 / 5203"
theme: journal
embed-resources: true
format: html
---

<div style="text-align: center; color: red; font-size: 48px; font-weight: bold; margin-top: 100px;">
ANSWER KEY
</div>

**Instructions:**
Please complete all components described below. In this assignment, you will work through a series of steps introducing you to working with data in *R*. You must address each part with either code, code and text, or just text. All necessary data is available in your `Data` folder under your Week 2 repository. 

**To submit:**

1. Clone your Week 2 Repo if you have not yet done so.
  * If you did this on Wednesday, hit `Pull` under the Git tab. This will being down the necessary data.

2. Create a new Quarto document. Title it "NAME_Lab2_Spring2026". 

3. At the very top of the document under title and author please be sure to put `embed-resources: true`. **This is a requirement each week!**

4. Complete all steps below.

5. Render and save. Make sure it is save in the same folder as the RProject file for Week 2. 

6. Commit and Push.

## Steps / Problems

### #1

Please import the penguins dataset. Assign it `penguin`.

```{r}
#| echo: true
#| eval: true

penguin <- read.csv('/Users/christopherwolfe/Library/CloudStorage/GoogleDrive-chriswolfe93091@gmail.com/My Drive/ECU_Courses/Spring2025/ANTH_Stat/data/penguins.csv')

#penguin <- read_csv('/Users/christopherwolfe/Library/CloudStorage/GoogleDrive-chriswolfe93091@gmail.com/My Drive/ECU_Courses/Spring2025/ANTH_Stat/data/penguins.csv')
```


### #2

How many rows and columns are in `penguin`?

```{r}
#| echo: true
#| eval: true

dim(penguin)
```


What type of data are in each column?

```{r}
#| echo: true
#| eval: true

# NOTE - str() can be used to complete both tasks
str(penguin)
```


### #3

Turn the `Sex` variable into a ***factor***. How many of each biological sex are there?

```{r}
#| echo: true
#| eval: true

penguin$sex <- as.factor(penguin$sex)

summary(penguin$sex)
```

Turn the `species` variable into a ***factor***. How many of each species are there?

```{r}
#| echo: true
#| eval: true

penguin$species <- as.factor(penguin$species)

summary(penguin$species)
```

### #4

Summarize the `penguin` dataset. What is the summary for `bill_depth` and `flipper_length`?



```{r}
#| echo: true
#| eval: true

summary(penguin)
# summary(penguin$bill_depth_mm)
# summary(penguin$flipper_length_mm)
```

### #5 

Create a vector of length 5 with the following names and assign it to new object called Hobbitses:

* Frodo, Samwise, Merry, Pippin, Bilbo

```{r}
#| echo: true
#| eval: true

Hobbitses <- c("Frodo", "Samwise", "Merry", "Pippin", "Bilbo")
Hobbitses
```

### #6 

Create a matrix with 2 rows and 3 columns. Fill the matrix will all numeric values.

```{r}
#| echo: true
#| eval: true

matrix(data = c(1:6), nrow = 2, ncol = 3)
```

### #7

Output the first 5 rows of `penguin`.

```{r}
#| echo: true
#| eval: true

head(penguin, 5)
```

Output the last 5 rows of `penguin`.

```{r}
#| echo: true
#| eval: true

tail(penguin, 5)
```

### #8

Select the value in the 2nd row and 3rd column of `penguin`.

What is this value?

```{r}
#| echo: true
#| eval: true

penguin[2,3]
```

### #9

Select rows 10-15 and columns 3-5 in `penguin`. 

Assign this selection as a new object and call it `sub_penguin`.

Turn this subset into a matrix.

```{r}
#| echo: true
#| eval: true

sub_penguin <- penguin[10:15, 3:5]

sub_penguin <- as.matrix(sub_penguin)
sub_penguin
```


### #10

Write me a sentence ***in code*** about penguins. 

*NB: Remember, data types and what is necessary to distinguish certain data type.*

```{r}
#| echo: true
#| eval: true

"Penguins do not live in the Arctic"
c("Penguins do not live in the Arctic")
c("Penguins", "do", "not", "live", "in", "the", "Arctic")

```

#

***Don't forget to save, render, commit, and push!***