---
title: "Lab 10 Regression Final"
subtitle: "ANTH 3900 / 5203"
format:
  html:
    toc: true
    toc-depth: 2
    toc-title: "Datasets"
    theme: cosmo
    code-fold: false
    embed-resources: true
execute:
  warning: false
  message: false
  eval: false
---

<span style="color: red;">**Note: If you want to use this qmd document to complete all analyses, you must change `eval: false` to `eval: true` in the header information.**</span>

\

***Due: Wednesday 4/15 by 10AM!***

Please complete all components described below.

**To submit:**

1. Clone your Week 13 Repo if you have not yet done so.

2. Create a new Quarto document. Title it "NAME_Lab10_Spring2026". 

3. At the very top of the document under title and author please be sure to put `embed-resources: true` AND `echo: true`. **This is a requirement each week!**

4. Complete all steps below.

5. Render and save. Make sure it is saved in the same folder as the RProject file for Week 13. 

6. Commit and Push.

## Instructions

For each of the five datasets below, you are given an outcome variable (y) and a set of predictors (x). **Your first task for every dataset is to identify the appropriate regression model from those covered in lecture and justify your choice in writing before running any code.** No hints are given about which model to use.

Your submission should be a rendered `.html` file knitted from this `.qmd` document. For each dataset you must provide:

1. A written paragraph identifying the regression type and your justification
2. All R code used to fit and evaluate the model
3. Written interpretation of results (in plain language where requested)

**Required packages:** Install these before beginning.

```{r}
#| label: setup

library(MASS)       # Pima.tr, glm.nb
library(AER)        # Affairs, NMES1988
library(DAAG)       # orings
library(margins)    # marginal effects for probit
library(tidyverse)  # optional but recommended
```

---

## Dataset 1: Pima Indians Diabetes

**Source:** `MASS::Pima.tr` — 200 Pima Indian women screened for Type II diabetes.

**Outcome (y):** `type` — whether the patient tested positive for diabetes

**Predictors (x):** `npreg`, `glu`, `bp`, `skin`, `bmi`, `ped`, `age`

### Step 1: Identify the model

*Replace this text with your written justification. What type of variable is `type`? What regression model is appropriate and why? (3–5 sentences)*

### Step 2: Fit the model

Fit a model predicting `type` from `glu`, `bmi`, `bp`, and `age`.

```{r}
#| label: dataset1-model

data("Pima.tr", package = "MASS")

# Examine the outcome variable
# YOUR CODE HERE

# Fit the model
fit1 <- # YOUR CODE HERE

summary(fit1)
```

### Step 3: Odds ratios

Exponentiate the coefficients to obtain odds ratios. Which predictors are statistically significant?

```{r}
#| label: dataset1-or

# Exponentiate coefficients
# YOUR CODE HERE
```

### Step 4: Written interpretation

*In a complete sentence written for a non-statistical audience, interpret the odds ratio for `glu`. What does a one-unit increase in glucose level mean for the odds of a diabetes diagnosis?*

---

## Dataset 2: Extramarital Affairs

**Source:** `AER::Affairs` — 601 respondents from a 1977 survey on extramarital affairs (Fair, 1978).

**Outcome (y):** `had_affair` — you will need to create this binary variable from `affairs`

**Predictors (x):** `gender`, `age`, `yearsmarried`, `children`, `religiousness`, `education`, `rating`

### Step 1: Create the outcome variable

```{r}
#| label: dataset2-setup

data("Affairs", package = "AER")

# Create binary outcome: 1 if affairs > 0, else 0
Affairs$had_affair <- # YOUR CODE HERE

# Check your recoding
table(Affairs$had_affair)
```

### Step 2: Identify the model and justify your choice

*Replace this text with your written justification. What regression model(s) could be applied to a binary outcome? Under what theoretical assumption would a probit link be preferred over a logit link? (4–6 sentences)*

### Step 3: Fit both models and compare

Fit both a logistic and a probit model, then compare AIC values.

```{r}
#| label: dataset2-models

# Logistic regression
fit2_logit <- # YOUR CODE HERE

# Probit regression
fit2_probit <- # YOUR CODE HERE

# Compare AIC
AIC(fit2_logit, fit2_probit)
```

### Step 4: Marginal effects

Using the `margins` package, compute average marginal effects for the probit model. Report the results for `rating` and `religiousness`.

```{r}
#| label: dataset2-margins

# Marginal effects
# YOUR CODE HERE
```

### Step 5: Written interpretation

*Interpret one marginal effect (either `rating` or `religiousness`) on the probability scale in plain language. What does this effect mean practically? (2–3 sentences)*

---

## Dataset 3: Warp Breaks

**Source:** `datasets::warpbreaks` — 54 observations of the number of warp breaks per loom when weaving wool yarn.

**Outcome (y):** `breaks` — count of warp thread breaks

**Predictors (x):** `wool` (type A or B), `tension` (Low, Medium, High)

### Step 1: Identify the model

*Replace this text with your written justification. What type of variable is `breaks`? Why is ordinary least squares (OLS) regression inappropriate here? What model is appropriate and why? (3–5 sentences)*

### Step 2: Fit the model

Fit a model including the interaction between `wool` and `tension`.

```{r}
#| label: dataset3-model

data("warpbreaks")

# Examine the outcome
# YOUR CODE HERE

# Fit model with interaction
fit3 <- # YOUR CODE HERE

summary(fit3)
```

### Step 3: Incidence rate ratios

Exponentiate the coefficients to obtain incidence rate ratios (IRR). What combination of wool type and tension level is associated with the most breaks?

```{r}
#| label: dataset3-irr

# Exponentiate coefficients
# YOUR CODE HERE
```

### Step 4: Written interpretation

*Summarize the model results. Which wool/tension combination produces the most breaks? Report the relevant IRR and interpret it in a complete sentence. (3–4 sentences)*

---

## Dataset 4: Physician Office Visits

**Source:** `AER::NMES1988` — 4,406 individuals from the 1987–88 National Medical Expenditure Survey.

**Outcome (y):** `visits` — number of physician office visits in the past year

**Predictors (x):** `hospital`, `health`, `chronic`, `gender`, `school`, `insurance`

### Step 1: Identify the model (initial)

*Replace this text with your initial thoughts. What type of variable is `visits`? What model would you reach for first? (2–3 sentences)*

### Step 2: Fit an initial model and check assumptions

```{r}
#| label: dataset4-poisson

data("NMES1988", package = "AER")

# Examine the outcome
# YOUR CODE HERE

# Fit initial model
fit4_initial <- # YOUR CODE HERE

summary(fit4_initial)
```

### Step 3: Reconsider and refit

Is the initial model appropriate? Fit a better-suited model.

```{r}
#| label: dataset4-refit

# Fit the improved model
fit4_final <- # YOUR CODE HERE

summary(fit4_final)

# Compare AIC
AIC(fit4_initial, fit4_final)
```

### Step 4: Interpret the IRR for `chronic`

```{r}
#| label: dataset4-irr

# Exponentiate coefficients from the final model
# YOUR CODE HERE
```

### Step 5: Written interpretation

*In 2–3 sentences, explain to a public health researcher why the final model was the right choice for this outcome. Then interpret the IRR for `chronic` in plain language.*

---

## Dataset 5: Space Shuttle O-Ring Failures

**Source:** `DAAG::orings` — data from 23 Space Shuttle launches prior to the Challenger disaster (January 28, 1986).

**Outcome (y):** `Total` (total O-ring incidents) out of 6 O-rings per launch. The dataset also contains `Erosion` and `Blowby` as component incident types.

**Predictors (x):** `Temperature`

### Step 1: Identify the model

*Replace this text with your written justification. How does this outcome differ from a standard binary (0/1) outcome? What regression model handles this structure and why? How does it differ from the binary logistic regression you used in Datasets 1 and 2? (4–6 sentences)*

### Step 2: Examine the data

```{r}
#| label: dataset5-setup

data("orings", package = "DAAG")

# Examine the data structure
head(orings)
str(orings)

# Plot Total incidents vs Temperature
plot(orings$Temperature, orings$Total / 6,
     xlab = "Temperature (°F)", ylab = "Proportion of O-rings with incidents",
     main = "O-ring incidents vs. launch temperature", pch = 19)
```

### Step 3: Fit the model

```{r}
#| label: dataset5-model

# Fit the logistic binomial model
# Hint: the outcome is specified as cbind(Total, 6 - Total)
fit5 <- # YOUR CODE HERE

summary(fit5)
exp(coef(fit5))
```

### Step 4: Predict the probability of failure at 31°F

On January 28, 1986, the temperature at launch was 31°F — far below any temperature in the training data.

```{r}
#| label: dataset5-prediction

# Create a new data frame for prediction
launch_day <- data.frame(Temperature = 31)

# Predict the probability of an O-ring incident (type = "response")
pred_prob <- predict(fit5, newdata = launch_day, type = "response")
pred_prob

# What is the probability that at least one of the 6 O-rings has an incident?
# YOUR CODE HERE
```

### Step 5: Written interpretation

*Write a 3–4 sentence summary of what this analysis tells us about the decision to launch the Challenger. Your summary should be written for a general audience with no statistical background. What does the predicted probability at 31°F suggest?*

---

## Submission Checklist

Before submitting, confirm:

- [ ] Each dataset section includes a **written justification** of the model choice *before* any code
- [ ] All code chunks run without errors in a clean R session
- [ ] Odds ratios / IRRs are reported and interpreted where requested
- [ ] All written interpretation questions are answered in full sentences
- [ ] The document is rendered to `.html` and submitted via the github portal
