---
title: "Lab 8 - LinearRegression_SP2026_KEY"
embed-resources: true
theme: journal
toc: true
format: html
editor_options: 
  chunk_output_type: console
---

**Please see below for answers to all parts of Lab 8. Note, I provide you one way to do the code - there are others. That said, all code relates to the topic of the week - linear regression.**

[**Scoring Summary**]{.underline}

* 10 points = submission to Github
  - **You must upload the QMD and HTML for full credit**
* 5 points = Code that runs (no errors, all chunks visible)
* 5 points = Correct answers. (Regardless of code)

# Answers

## Exercise 1

```{r}
#| echo: true
#| message: false
library(tidyverse)
penguins <- penguins
```

### Step 1

```{r}
#| echo: true

penguins %>% ggplot(aes(flipper_len, body_mass)) + geom_point() + geom_smooth(method = "lm", se=F) + theme_minimal()
```

### Step 2

```{r}
#| echo: true

model_1 <- lm(body_mass ~ flipper_len, data = penguins)
summary(model_1)
```

### Step 3

$\beta_1$ is equal to 49.686 grams. This means that for every 1 unit increase in flipper lenght (millimeters), body mass increases by 49.686 grams.

### Step 4

The multiple $R^2$ is 0.759. The adjust $R^2$ is 0.7583. This means 75% of the variation in body mass can be explained by flipper length. 

## Exercise 2

```{r}
#| echo: true

cars <- mtcars
```

### Step 1

```{r}
#| echo: true

model_2 <- lm(mpg ~ wt + hp, data = cars)
summary(model_2)
```

### Step 2

Holding the weight variable constant, for every 1 unit increase in horsepower, there is a decrease in miles per gallon by approximately 0.03 miles. 

### Step 3

Generally, the error or variance is relative homogenous / homoscedastic. 

```{r}
#| echo: true

# plot(model_2) OR see below

# Extract fitted values and residuals
model_2_diag <- data.frame(
  fitted    = fitted(model_2),
  residuals = residuals(model_2)
)

ggplot(model_2_diag, aes(x = fitted, y = residuals)) +
  geom_point(shape = 21, fill = "steelblue", color = "black", size = 3, alpha = 0.8) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red", linewidth = 0.8) +
  labs(
    title = "Residual Plot",
    subtitle = "Model: mpg ~ wt + hp",
    x = "Fitted Values",
    y = "Residuals"
  ) +
  theme_classic(base_size = 14) +
  theme(
    plot.title    = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, color = "grey40"),
    axis.text     = element_text(color = "black")
  )
```

### Step 4

Note - 0 hp or 0 wt does not make sense. Common practice is to set equal to a mean value when holding 'constant.'

```{r}
#| echo: true

# 1. Predicted MPG for a car with 500 hp (using mean wt as reference)
predict(model_2, newdata = data.frame(wt = mean(mtcars$wt), hp = 500))

# 2. Predicted MPG for a car weighing 7.0 (using mean hp as reference)
predict(model_2, newdata = data.frame(wt = 7.0, hp = mean(mtcars$hp)))

# 3. Predicted MPG for a car with 450 hp and weight 6.5
predict(model_2, newdata = data.frame(wt = 6.5, hp = 450))
```

## Exercise 3

### Step 1

```{r}
#| echo: true

penguins %>% ggplot(aes(flipper_len, body_mass, color = species)) + geom_point() + geom_smooth(method = "lm", se=F) + theme_minimal()
```

### Step 2

```{r}
#| echo: true

model_3 <- lm(body_mass ~ flipper_len + species, data = penguins)
summary(model_3)
```

### Step 3

Because *Adelie* is the reference species, if we hold flipper length constant, then *Gentoo* body mass is 266.810 grams greater than *Adelie* body mass. Therefore, the expected difference is 266.810.

```{r}
#| echo: true

# Adelie
adelie <- -4031.477 + 266.810 + 40.705

# Gentoo
gentoo <- -4031.477 + 40.705

adelie - gentoo
```

