---
title: "Lab 8 - Linear Regression"
subtitle: "ANTH 3900 / 5203"
theme: journal
embed-resources: true
echo: true
format: html
editor_options: 
  chunk_output_type: console
---

***Due: Monday 3/30 by end of day***

Please complete all components described below.

**To submit:**

1. Clone your Week 11 Repo if you have not yet done so.

2. Create a new Quarto document. Title it "NAME_Lab8_Spring2026". 

3. At the very top of the document under title and author please be sure to put `embed-resources: true` AND `echo: true`. **This is a requirement each week!**

4. Complete all steps below.

5. Render and save. Make sure it is saved in the same folder as the RProject file for Week 11. 

6. Commit and Push.

**Remember:**

1.  Whenever you are using a package. You must open it. Remember `library()`. This must be in a code chunk in your document. 
2.  If you are using data, you must import it inside a code chunk. You cannot just use the import data button in the `Environment` pane. 
3.  Pay attention to your warnings and errors. They mean something.
  - Remember, ggplot requires a `+` sign, not the pipe.
4. Remember to `set.seed()` if necessary.

## Exercise 1: The Baseline

```{r}
#| echo: true

penguins <- penguins
```


Biologists want to estimate the weight of a penguin just by measuring its flipper. You will build a simple linear model to see if this is a reliable method.

The equation you are estimating is:
$$
BodyMass=\beta_0+\beta_1(FlipperLength)+\epsilon
$$

### Step 1: Visualization

Create a scatterplot mapping flipper_len to the x-axis and body_mass to the y-axis from the penguins dataset. Add a linear trend line.

### Step 2: Modeling

Fit the linear model using lm() and save it as an object called model_1. Then, [display]{.underline} the summary of the results.

### Step 3: Interpretation 

What is the value of the slope ($\beta_1$)? Explain what this number means in plain English, using the correct units (millimeters and grams).

### Step 4: Model Fit 

Report the Multiple $R^2$ value. Write one sentence explaining what the value is telling us?

## Exercise 2: Multiple Regression

```{r}
#| echo: true

cars <- mtcars
```


A car buyer wants to know how engine power and vehicle weight affect fuel efficiency. You will build a multiple regression model to parse out these overlapping effects using the mtcars dataset.

The equation you are estimating is:
$$
MPG=\beta_0+\beta_1(Weight)+\beta_2(Horsepower)+\epsilon
$$

### Step 1: Modeling

Fit a multiple regression model predicting mpg (miles per gallon) using wt (weight in 1000 lbs) and hp (horsepower). Save it as model_2 and display the results.

### Step 2: The "Partial" Interpretation 

Question: Look at the coefficient for horsepower ($\beta_2$). Explain what this specific number means while explicitly mentioning what happens with the weight variable.

### Step 3: Diagnostics

Generate a residuals vs. fitted plot for your model.(Hint: Use `plot(model_2)` OR use `resid(model_2)` and plot by hand using ggplot). Interpret your plot.

### Step 4: Predictions

1.  What is the predicted MPG of a car with 500 hp? (AKA holding wt constant)

2.  What is the predicted MPG of a car weighing 7.0 (7000 lbs).

3.  What is the predicted MPG of a car with 450 HP and weight 6.5 thousand lbs?

## Exercise 3: Categorical Predictors

We know flipper length predicts weight, but does a Gentoo penguin weigh the same as an Adelie penguin if they have the exact same flipper length? Let's add a categorical variable (text) to our model.

The equation you are estimating is:
$BodyMass=\beta_0+\beta_1(FlipperLength)+\beta_2(Species)+\epsilon$

### Step 1: The Visual Setup

Create the same scatterplot as you did in Exercise 1, but this time, color the points by species.

### Step 2: Modeling

Fit the multiple regression model predicting body_mass with both flipper_len and species. Save it as model_3 and display and interpret the results.

### Step 3: Interpretation

Question: Interpret the coefficient for the Gentoo species. If an Adelie and a Gentoo have the exact same flipper length, what is the expected difference in their body mass?

  * Hint: Write out the regression equation for each species and then find the difference(s).
  
#

***Don't forget to save, render, commit, and push!***