---
title: "Midterm Practical - Spring 2026 - KEY"
embed-resources: true
theme: journal
toc: true
format: html
editor_options: 
  chunk_output_type: console
---

**Please see below for answers to all parts of the Midterm. Note, I provide you one way to do the code - there are others. You must demonstrate you have the correct answer for full credit.**

# Answers

## Section 1

```{r}
#| echo: true
#| message: false

library(tidyverse)
svad <- read.csv("data/svad_lb.csv")
```


### Question 1

```{r}
#| echo: true

# Inspect the dataset

glimpse(svad)
```

```{r}
#| echo: true

# Replace all -1 values with NA
svad_clean <- svad %>%
  mutate(across(where(is.numeric), ~ na_if(.x, -1)))

# Check
svad_clean %>%
  summarise(across(where(is.numeric), ~ any(.x == -1, na.rm = TRUE)))
```

```{r}
#| echo: true

# Filter for female individuals and count missing values in measurement columns
# Adjust 'sex' and female indicator based on actual column name/coding in dataset
female_missing <- svad_clean %>%
  filter(SEX == "F") %>%  
  select(where(is.numeric)) %>%
  summarise(across(everything(), ~ sum(is.na(.x)))) %>%
  pivot_longer(everything(),
               names_to  = "column",
               values_to = "missing_count")
```

```{r}
#| echo: true

female_missing
```

### Question 2

```{r}
#| echo: true

crural <- svad_clean %>%
  filter(agey >= 0 & agey <= 5,
         FDL_L >= 90) %>%
  select(contains(c("FDL", "TDL", "FBDL", "HDL", "RDL", "UDL"))) %>%
  mutate(crural = TDL_L / FDL_L)

head(crural, 3)
```

### Question 3

```{r}
#| echo: true

group1 <- svad_clean %>% filter(SEX != "NA") %>%
  group_by(SEX) %>%
  summarise(
    across(
      contains(c("FDL", "TDL", "FBDL", "HDL", "RDL", "UDL")),
      list(
        n   = ~ sum(!is.na(.x)),
        mean = ~ mean(.x, na.rm = TRUE),
        sd   = ~ sd(.x, na.rm = TRUE)
      )
    )
  ) %>%
  pivot_longer(
    cols      = -SEX,
    names_to  = c("measurement", "stat"),
    names_sep = "_(?=[^_]+$)",  # splits on the LAST underscore
    values_to = "value"
  )

head(group1)
```

### Question 4

```{r}
#| echo: true

group2 <- svad_clean %>%
  mutate(age_int = as.integer(agey)) %>%
  group_by(age_int) %>%
  summarise(
    across(
      contains(c("FMSB", "FDB", "TPB", "TMSB", "TDB")),
      list(
        n    = ~ sum(!is.na(.x)),
        mean = ~ mean(.x, na.rm = TRUE),
        sd   = ~ sd(.x, na.rm = TRUE)
      )
    )
  ) %>%  pivot_longer(
    cols      = -age_int,
    names_to  = c("measurement", "stat"),
    names_sep = "_(?=[^_]+$)",  # splits on the LAST underscore
    values_to = "value"
  )

head(group2)
```

### Question 5

```{r}
#| echo: true

svad_clean <- svad_clean %>%
  rename_with(~ str_remove(.x, "_L$") %>% str_to_lower(),
              ends_with("_L"))

# Print column names to verify
names(svad_clean)
```

## Section 2

```{r}
#| echo: true
#| message: false

palm <- read.csv("data/palm.csv")
head(palm)
```

### Question 6

```{r}
#| echo: true

palm_clean <- palm %>% na.omit()

ggplot(palm_clean, aes(x = stem_solitary, y = max_stem_dia_cm, fill = stem_solitary)) +
  geom_violin(alpha = 0.4) +
  scale_fill_manual(values = c("non-solitary" = "steelblue", "solitary" = "tomato", "both" = "goldenrod")) +
  labs(
    title    = "Maximum Stem Diameter by Stem Growth Form",
    subtitle = "Comparison of solitary and clustered palm species",
    x        = "Stem Growth Form",
    y        = "Maximum Stem Diameter (cm)",
    fill     = "Growth Form"
  ) +
  theme_classic(base_size = 14) +
  theme(
    plot.title      = element_text(face = "bold", hjust = 0.5),
    plot.subtitle   = element_text(hjust = 0.5, color = "grey40"),
    legend.position = "none",
    axis.line       = element_line(color = "black"),
    axis.text       = element_text(color = "black")
  )
```

### Question 7

There is a linear relationship between stem height and diameter with a taller stem generally associated with a larger diameter.

```{r}
#| echo: true

palm_clean %>% ggplot(aes(max_stem_height_m, max_stem_dia_cm, colour = stem_solitary)) + geom_point() + geom_smooth(se=F, method = "lm") + theme_minimal()
```

### Question 8

Could be anything within reason.

```{r}
#| echo: true

palm_clean %>% ggplot(aes(average_fruit_length_cm, average_fruit_width_cm, colour = fruit_size_categorical)) + geom_point() + geom_smooth(se=F, method = "lm") + theme_minimal()
```

## Section 3

```{r}
#| echo: true
#| message: false

diwali <- read.csv("data/diwali.csv")
head(diwali)
```

### Question 9

#### Normality

Data are non-normal.

```{r}
#| echo: true

qqnorm(diwali$Amount, frame = F)
qqline(diwali$Amount, col = "steelblue", lwd = 3)
```

```{r}
#| echo: true

ggplot(diwali, aes(x = Amount)) + theme_bw() + geom_histogram(aes(y = after_stat(density)), color = "black", fill = "white") + geom_density(fill = "steelblue", alpha = 0.4) + facet_grid(cols = vars(Zone))
```

#### Homoscedasticity

Variance is NOT homoskedastic.

```{r}
#| echo: true

car::leveneTest(Amount ~ Zone, data = diwali)
```

### Question 10

While this may show "outliers", I see no reason to remove given they could be realistic.

```{r}
#| echo: true

diwali %>% ggplot(aes(x = Zone, y = Amount)) + theme_bw() + geom_boxplot(outlier.colour = "red", outlier.size = 3)
```

### Question 11

The data are non-normal with more than 2 groups. Therefore, I use the kruskal wallis test.

There are statistically significant differences in spending by zone across *most* regional comparisons. 

```{r}
#| echo: true

kruskal.test(Amount ~ Zone, data = diwali)
```

```{r}
#| echo: true

library(dunn.test)

print(dunn.test(diwali$Amount, diwali$Zone, method = "holm"))
```


### Question 12

The data are non-normal. The variance is homoscedastic. No outliers were removed. A man-whitney test is run because of the binary category.

There are not statistically significant differences by marital status.

```{r}
#| echo: true

ggplot(diwali, aes(x = Amount)) + theme_bw() + geom_histogram(aes(y = after_stat(density)), color = "black", fill = "white") + geom_density(fill = "steelblue", alpha = 0.4) + facet_grid(cols = vars(Marital_Status))
```

```{r}
#| echo: true

car::leveneTest(Amount ~ as.factor(Marital_Status), data = diwali)
```

```{r}
#| echo: true

diwali %>% ggplot(aes(x = as.factor(Marital_Status), y = Amount)) + theme_bw() + geom_boxplot(outlier.colour = "red", outlier.size = 3)
```

```{r}
#| echo: true

wilcox.test(Amount ~ Marital_Status, data = diwali)
```

## Section 4

```{r}
#| echo: true

penguin <- penguins
```

### Question 13

```{r}
#| echo: true

min_max <- function(body_mass) {
  minimum <- sort(body_mass, na.last = NA)[1]
  maximum <- sort(body_mass, na.last = NA)[length(sort(body_mass, na.last = NA))]
  
  return(c(minimum = minimum, maximum = maximum))
}

# Example usage
min_max(penguin$body_mass)

# Check
min(penguin$body_mass, na.rm = T)
max(penguin$body_mass, na.rm = T)
```

### Question 14

```{r}
#| echo: true

# Define the columns to iterate over
cols <- c("bill_len", "bill_dep", "flipper_len", "body_mass")

# Empty data frame to store results
results <- data.frame(
  variable = character(),
  minimum  = numeric(),
  maximum  = numeric()
)

# For loop
for (col in cols) {
  vals    <- min_max(penguins[[col]])
  results <- rbind(results, data.frame(
    variable = col,
    minimum  = vals["minimum"],
    maximum  = vals["maximum"],
    row.names = NULL
  ))
}

print(results)
```

